---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# DISCLAIMER

This package/application is provided without warranty of any kind.

While every effort has been made to ensure that that routines provide correct and accurate results, all responsibility for the use of the results belongs to the user.


# PooledInfRate

The PooledInfRate package provides functions to estimate proportions from pooled or group testing data.

## Installation

You can install PooledInfRate from GitLab:

``` r
# FILL THIS IN! HOW CAN PEOPLE INSTALL YOUR DEV PACKAGE?
devtools::install_git("https://git.cdc.gov/bkb5/PooledInfRate")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(PooledInfRate)
## basic example code
x <- c(1,0,0,0)
m <- c(50,25,10,5)
pooledBin(x,m)
```
